Community-Centric Innovation

### 2014-03-01 paper.tex
changed abstract

\author{
	Name~Surname,~%\IEEEmembership{Member,~IEEE,}
    Name~Surname,~%\IEEEmembership{Member,~IEEE,}
    Name~Surname,~%\IEEEmembership{Member,~IEEE,}
    Name~Surname,~%\IEEEmembership{Member,~IEEE,}
    and~Name~Surname%~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
\thanks{The authors are with\newline
blah \newline
blah \newline
blah}%
}

@@

\begin{abstract}
%\boldmath
The Internet is profoundly transforming the economic and social structures of the 21st century.
The ability to instantaneously share data across the globe enables new ways of collaboration and production models.
The individuals can join loosely organized working groups and communities and contribute to the evolution and progress of the humankind.
By choosing which communities they join and which communities they contribute to, the users can also be participants and producers.
Furthermore, this model enables a new way of democracy in which the people vote not only with their ballots, but also with their daily jobs and contributions.
In this work we describe examples of bottom-up initiatives in which the users play a central role in the development of the project.
These are initiatives by and for the people.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

@@

In this work we consider several examples in which the users organize in communities to construct user-centric networks and networked services.

\section{Community Networks}
Probably the best example of user-centric networking are community networks.
Community networks are deployed by the same users that use them.
Consequently, they are carefully tailored to satisfy the users' needs.
This needs are not necessarily only about connectivity.
The users might be interested in learning about the network, and in researching  new services, applications and protocols.

@@

The decentralized nature of community networks greatly helps in fostering organic growth and also in  resisting node failures.
The absence of central points of failure guarantees that two neighbours will be able to connect to each other as long as they keep their nodes running, regardless of what happens to the rest of the network.
Mesh routing protocols, such as BMX6, allow the network to quickly re-configure itself and to route packets around failures to keep connectivity.
These protocols are particularly tailored to work in challenging multi-hop wireless conditions.
To take down a community network it is necessary to shut down all the nodes.

@@

A core principle of community networks is precisely the sharing of the network as a commons good.
There is no single owner of the network.
%In fact there is no owner of the network.
Even though the participants may own the pieces of hardware that they contributed to the network, the network as a whole belongs the community.
As the value of the network increases with the number of participants, the community has an incentive to share existing network resources with newcomers.

@@

\section{Networked Cryptocurrencies}

@@

Peer-to-peer networks are extremely resilient.
There is no single point of failure or central entity that can be attacked to take down the network.
As there are thousands of nodes spread across the network, the currency can survive as long as the underlying data network survives.
The risks of blockades or political interference are greatly mitigated.
Many Internet organizations such as Wikileaks and Wordpress accept donations and payments in cryptocurrency.
In some cases to bypass governments blockades and in others to operate in countries with lacking payment networks.

@@

\section{Community-Centric Innovation}

@@

\section{Projects about Community Networks and Cryptocurrencies}
Community-centric innovation such as Community Networks and Networked Cryptocurrencies have attracted the attention of the research community.
In this section we detail three European projects that are related to community-centric innovation.

@@

\section{Conclusions}
In some aspects of information and communication technologies, communities are  leading the path of innovation.
Community-centric innovation is characterized by peer-to-peer relations and openness.
Communities have also demonstrated their ability to take advantage of data networks and network effects.
In the case of community networks, a reinforcement loop is established in which a community deploys a network that, in its turn, strengthens the community.
Community networks have been with us for over a decade and have remained faithful to the original spirit of the Internet in terms of decentralization.
These networks are planned, financed, deployed and maintained by the same people that uses them.
They adopt self-management dynamics and can benefit from the combined knowledge of all the members of the community.

>>>>

\author{
    Jaume~Barcelo~%\IEEEmembership{Member,~IEEE,}
    and~Boris~Bellalta%~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
\thanks{The authors are with Universitat Pompeu Fabra.\newline
Address: Roc Boronat 138, 08018, Barcelona. \newline
E-mail: {name.surname}@upf.edu }%
}

@@

\begin{abstract}
%\boldmath
The Internet is profoundly transforming the economic and social structures of the 21st century.
The ability to instantaneously share data across the globe enables new ways of collaboration and new production models.
The individuals can join loosely organized working groups and communities and contribute to the evolution and progress of the humankind.
By choosing which communities they join and which communities they contribute to, the users can also be participants and producers.
%Furthermore, this model enables a new way of democracy in which the people vote not only with their ballots, but also with their daily jobs and contributions.
In this work we describe examples of community-centric innovation in which the users play a central role in the development of the project.
The emphasis is placed in community networks and networked cryptocurrencies.
These technologies are good examples in the sense that they empower the communities that use and develop them.
These are initiatives by and for the people.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

@@

In this work we consider several examples in which the users organize in communities to construct user-centric networks and networked services.
We start with community networks in Sec.~\ref{sec:community-networks} and then continue with networked cryptocurrencies in Sec.~\ref{sec:networked-cryptocurrencies}.
Other technologies fruit of community-centric innovation are outlined in Sec.~\ref{sec:other} and an overview of research and innovation projects related to community-centric innovation is presented in Sec.~\ref{sec:projects}.
Finally, we conclude in Sec.~\ref{sec:conclusion}.

\section{Community Networks}
\label{sec:community-networks}

Probably the best example of user-centric networking are community networks.
Community networks are deployed by the same users that use them.
Consequently, they are carefully tailored to satisfy the users' needs.
This needs are not necessarily only about connectivity.
The users might be interested in learning about the network, and in researching  new services, applications and protocols.

@@

The decentralized nature of community networks greatly helps in fostering organic growth and also in  resisting node failures.
The absence of central points of failure guarantees that two neighbours will be able to connect to each other as long as they keep their nodes running, regardless of what happens to the rest of the network.
Mesh routing protocols, such as BMX6 \cite{neumann2012ebc}, allow the network to quickly re-configure itself and to route packets around failures to preserve connectivity.
These protocols are particularly tailored to work in challenging multi-hop wireless conditions.
To take down a community network it is necessary to shut down all the nodes.

@@

A core principle of community networks is precisely the sharing of the network as a commons good.
This is particularly relevant in the case of wireless networks that use the ISM (industrial, scientific, medical) bands.
This bands contain a small fraction of the spectrum that is available without requiring a licence. 
In practice, these bands are a commons good that should be carefully managed by the community to use it effectively without creating unnecessary interference.
Good practices include the use of directive antennas and the limitation of the radiated power.

As the community network is a commons resource, there is no single owner of the network.
%In fact there is no owner of the network.
Even though the participants may own the pieces of hardware that they contributed to the network, the network as a whole belongs the community.
As the value of the network increases with the number of participants, the community has an incentive to share existing network resources with newcomers.

@@

\section{Networked Cryptocurrencies}
\label{sec:networked-cryptocurrencies}

@@

Peer-to-peer networks are extremely resilient.
There is no single point of failure or central entity that can be attacked to take down the network.
As there are thousands of nodes spread across the network, the currency can survive as long as the underlying data network survives.
The risks of blockades or political interference are greatly mitigated.
Many Internet organizations such as Wikileaks and Wordpress accept donations and payments in cryptocurrency.
In some cases the reason to bypass government or paying networks blockades and in others the motivation is to operate in countries with lacking payment networks.

It is in areas lacking effective financial and banking systems where cryptocurrencies have a potential for greater impact.
Just like some areas leapfrogged from no telecommunication services to wireless handheld Internet enabled devices, it is possible that these same areas that currently use mobile calls minutes as a currency transition directly to the latest currency technology which is cryptocurrency.
Many countries suffer from chronic hyperinflation problems that could be solved by moving to a cryptocurrency in which the inflation rate (if any) can be chosen by the community and embedded in the protocol.

@@

\section{Community-Centric Innovation}
\label{sec:other}

@@

\section{Projects about Community Networks and Cryptocurrencies}
\label{sec:projects}

Community-centric innovation such as Community Networks and Networked Cryptocurrencies have attracted the attention of the research community.
In this section we detail three European projects that are related to community-centric innovation.

@@

\section{Conclusions}
\label{sec:conclusion}

In some aspects of information and communication technologies, communities are  leading the path of innovation.
Community-centric innovation is characterized by peer-to-peer relations and openness.
Communities have also demonstrated their ability to take advantage of data networks and network effects.
In the case of community networks, a reinforcement loop is established in which a community deploys a network that, in its turn, strengthens the community.
Community networks have been with us for over a decade and have remained faithful to the original spirit of the Internet in terms of decentralization.
These networks are planned, financed, deployed and maintained by the same people that uses them.
They adopt self-management dynamics and can benefit from the combined knowledge of all the members of the community.
